---
title: "Bycicle Project"
output: html_notebook
---

This is a project for the Google Data Analyst certificate on Coursera.org. The

```{r}

#load libraries and Datasets
library(tidyverse)
library(dplyr)
library(readr)
install.packages("geosphere")
library(geosphere)

# Using the Q1 from 2020 available on <https://divvy-tripdata.s3.amazonaws.com/index.html>
Divvy_Trips_2020_Q1 <- read_csv("DATA/Divvy_Trips_2020_Q1.csv")


# renaming the dataset for improve handling
DF_20 <- Divvy_Trips_2020_Q1

DF_20 <- DF_20 %>% 
  rename(user_type = member_casual)


# FILTERING OUT THE TRIPS WITH NON SIGNIFICANT DURATION (less than 60 seconds)
DF_20 <- DF_20 %>% mutate(trip_duration = ended_at - started_at)
DF_20_filtered <- DF_20 %>% filter(trip_duration >= 60) 

# CALCULATING TRIP DISTANCE

DF_20_filtered <- DF_20_filtered %>% rowwise() %>% mutate(trip_station_distance = distm (c(start_lat, start_lng), c(end_lat, end_lng), fun = distHaversine) )

```

CREATING SUBSETS for further analysis

```{r}

Casual_subset <- DF_20_filtered %>% filter(user_type == "casual")

member_subset <- DF_20_filtered %>% filter(user_type == "member")

round_trips_subset <- DF_20_filtered %>%  filter(trip_station_distance == 0)

inter_station_trip_subset <- DF_20_filtered %>%  filter(trip_station_distance >= 1)
```

EXTRACTING RELEVANT INFORMATION 
\### ROUND TRIPS \###

```{r}
            
sum(Casual_subset$trip_station_distance == 0, rm.na=TRUE )
# 6489
mean(round_trips_subset$trip_duration[round_trips_subset$user_type == "casual"])
# Time difference of 4571.013 secs ~ about 1h16min

###

sum(member_subset$trip_station_distance == 0, rm.na=TRUE )
# 6258
mean(round_trips_subset$trip_duration[round_trips_subset$user_type == "member"])
# Time difference of 1402.789 secs ~ about 23min

```

### INTER STATION TRIPS ###

```{r}
sum(Casual_subset$trip_station_distance >= 1, rm.na=TRUE )
# 38042
mean(inter_station_trip_subset$trip_duration[inter_station_trip_subset$user_type == "casual"])
#Time difference of 6544.039 secs ~ about 1h49min

###

sum(member_subset$trip_station_distance >= 1, rm.na=TRUE )
# 368431
mean(inter_station_trip_subset$trip_duration[inter_station_trip_subset$user_type == "member"])
#Time difference of 757.1886 secs ~ about 12 min
```

### GRAPH TIME ###

```{r}

 df_proporciones <- DF_20_filtered %>%
     count(user_type) %>%
    mutate(proporcion = round (100 * (as.numeric(n) / as.numeric(sum(df_proporciones$n))),2)) 
  



 # Crear el gráfico de pastel
 ggplot(df_proporciones, aes(x = "", y = proporcion, fill = user_type)) +
    geom_bar(width = 1, stat = "identity") +
    coord_polar("y", start = 0) +
    theme_void() +
    scale_fill_brewer(palette = "Set1") + # Colores personalizados
    labs(title = "Proporción de Tipos de Usuarios", fill = "Tipo de Usuario") + theme(plot.title = element_text(hjust = 0.5)) +
    geom_text(aes(label = paste0(round(proporcion, 2), "%")), position = position_stack(vjust = 0.5))
```



Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
